### 2015-11-10 postdoctoral/S_prime_kappa.tex
start scaffolding results with Alan

  \begin{remark}
    Scheeper's \(S(\kappa)\) requiring injections is stronger than my
    \(S'(\kappa)\) requiring finite-to-one maps. Dow suggests that
    \(S'(\omega_\omega)\) holds in ZFC by the following.
  \end{remark}

@@

  \begin{theorem}
    For each \(k<\omega\) there exists a topology on \(\omega_k\) which
    is \(\omega\)-bounded and locally countable.
  \end{theorem}

  \begin{proof}
    TODO: write up Alan's proof
  \end{proof}

  \begin{definition}
    A Kurepa family \(\mc K\)
    is a family of countable sets such that for each \(A\in\mc K\),
    \(\mc K\rest A=\{K\cap A:K\in\mc K\}\) is countable.
  \end{definition}

@@

  \begin{proof}
    This is actually a corollary of an observation of Todorcevic communicated
    by Dow in [TODO cite Gen Prog in Top I]:
    if every Kurepa family of size at most \(\theta\)
    extends to a cofinal Kurepa family, then the same is true of \(\theta^+\).
    So the result follows as
    every Kurepa family of size \(\omega\) extends to
    the cofinal Kurepa family \([\omega]^\omega\). (TODO am I understanding
    this right?)

    We may alternatively
    obtain the result from the previous topological argument by using the family
    \(\mc K\) of compact sets in the constructed topology on
    \(\omega_k\) as our witness. Of course, every Lindel\"of set in
    a locally countable space is countable. Thus \(\mc K\)
    is cofinal since for every countable set \(A\), \(\cl A\) is compact
    and countable.
    It is Kurepa since for every compact (countable) set \(A\), (TODO).
  \end{proof}

  \begin{theorem}
    \(S'(\omega_k)\) for each \(k<\omega\).
  \end{theorem}

  \begin{proof}
    Let \(k<\omega\), and \(\mc K\) be a cofinal Kurepa family on \(\omega_k\).
    (TODO finish)
  \end{proof}

>>>>

  (joint work with Alan Dow)

  \begin{definition}
    Two functions \(f,g\) are almost compatible if
    \(\{a\in\dom f\cap\dom g:f(a)\not=g(a)\}\) is finite.
  \end{definition}

  \begin{definition}
    \(S'(\theta)\) states that there exists a cofinal family
    \(\mc S\subseteq[\theta]^{\omega}\) and a collection of pairwise
    almost compatible
    finite-to-one functions \(\{f_S\in\omega^S:S\in\mc S\}\)
  \end{definition}

  \begin{definition}
    \(S(\theta)\) strengthens \(S'(\theta)\) by requiring the collection
    to contain one-to-one functions.
  \end{definition}

  We wish to show that Scheeper's original \(S(\theta)\) is strictly stronger
  than \(S'(\theta)\).

@@

  % TODO: read through Alan's proof and fit to rest of document
          \begin{theorem}
           For each $n\in \omega$, there is a locally countable,
            $\omega$-bounded topology on $\omega_n$. Note
            that this means that the closure of any set has
            the same cardinality and weight as the set.
          \end{theorem}

          To prove the theorem, we must actually prove a stronger lemma.

          \begin{lemma}
          Assume that $X$ has cardinality at most $\omega_n$ (for
          any $n\in \omega$),
          and is locally countable, locally compact, and the closure
          of each set has the same cardinality as the set.
          Then $X$ has an $\omega$-bounded extension with the
          same properties.
          \end{lemma}

            \begin{proof}
            We prove this by induction on $n$.
            In fact, we make our inductive statement that  if $\tilde X$
            is the extension of $X$, then  $\tilde X \setminus X$ also
            has cardinality $\omega_n$.
             If $n=0$,
              then we can just take the free union of two copies of $X$
              and then the one-point compactification.
             So suppose $n>0$ and
            that $X$ is such a topology on the ordinal $\omega_n$.
            For each $\alpha<\omega_n$, the closure of the initial
            segment $\alpha$ is bounded by some $\gamma_\alpha$.
            Also, because $X$ is locally countable, $\gamma_\alpha$
            can be chosen so that $\alpha$ is contained  in the interior
            of $\gamma_\alpha$. There is a cub  $C\subset \omega_n$
            with the property that for each $\delta\in C$ and $\alpha<\delta$,
             $\gamma_\alpha$ is also less than $\delta$. This implies
             that for each $\delta\in C$, the initial segment $\delta$ is
             open, and if $\delta$ has uncountable cofinality, then
              $\delta$ is clopen.

              The proof will be easier to visualize if we now
              identify the points of $X$ with the point set
               $\omega_n\times \{0\}$ and we will add the points
                $\omega_n\times \{1\}$ to create the extension.
             By induction on $\lambda \in C$ we define a topology
             on $\omega_n\times\{0\}\cup \lambda\times \{1\}$ so that
             $\omega_n \times \{0\}$ is an open subset. We also ensure,
             by induction, for each $\alpha<\lambda$,
             the closure of $\alpha\times 2$ is an
              $\omega$-bounded  subset of $\lambda\times 2$.

          In the case that $n=1$, then  choose any sequence
           $\lambda_n : n\in \omega$ increasing cofinal in $\lambda$.
           If $\lambda $ is a limit in $C$, then we simply take the
           topology we have constructed so far on $\lambda\times 2$
           and there's nothing more needs to be done. Otherwise
          we may assume that $\lambda_0$ is the predecessor
          of $\lambda\in C$ and
           we set $Y_\lambda $ to equal the countable set
            $\overline{\lambda}\setminus \lambda$. For convenience,
             and with no loss, we assume that $\lambda$ itself is a limit
             of limits.
           And we have a topology on
           $$\lambda_0\times 2 \cup (\lambda \cup Y_\lambda)\times\{0\}~.$$
            Recursively
            choose clopen sets $U_n$ in this topology
             so that $\lambda_0\times 2\subset U_0$,
             $U_{n}\cup \lambda_{n+1}\times \{0\}$ is contained
              $ U_{n+1}$ while $U_{n+1}$ is disjoint from $Y_\lambda$.
              It is easy to see that we can have all the points
              in $(\lambda\setminus \{\lambda_n : n\in\omega\})\times \{1\}$
              be isolated, and arrange that $(\lambda_n,1) $ is the
              point at infinity in the one-point compactification
              $U_n\cup (\lambda_n\times \{1\})$.

                Now we handle the case $n>1$ and we can shrink $C$
                and now assume
                that $C$ is the closure of $\{\lambda\in C : \mbox{cf}(\lambda)>\omega\}$.
                We again proceed by induction on $\lambda\in C$.  If
                $\lambda$ is a limit in $C$, then there is nothing to do: we
                simply have defined an appropriate topology on
                $\omega_n\times  \{0\}\cup \lambda\times \{1\}$ so that
                for each $\mu\in C\cap \lambda$ with $\mbox{cf}(\mu)>\omega$,
                 $\mu\times 2$ is a clopen $\omega$-bounded subspace.
                In case $\lambda$ is not a limit of $C$, then $\lambda$
                has uncountable cofinality and a predecessor $\mu\in C$.
                We therefore have that $\lambda\times\{0\}$ is clopen
                in $\omega_n\times \{0\}$.
                 We apply the induction hypothesis
                to the space $\lambda\times \{0\} \cup \mu\times 2$ to choose
                the topology on $\lambda\times 2$.

            \end{proof}

  \begin{definition}
    A Kurepa family \(\mc K\subseteq[\theta]^{\omega}\) on \(\theta\)
    satisfies that
    \(\mc K\rest A=\{K\cap A:K\in\mc K\}\) is countable
    for each \(A\in[\theta]^\omega\).
  \end{definition}

@@

  \begin{proof}
    This is actually a corollary of an observation of Todorcevic communicated
    by Dow in [TODO cite Gen Prog in Top I]:
    if every Kurepa family of size at most \(\theta\)
    extends to a cofinal Kurepa family, then the same is true of \(\theta^+\).
    So the result follows as
    every Kurepa family \(\mc K\) of size \(\omega\) extends to
    the cofinal Kurepa family \([\bigcup\mc K]^\omega\).

    We may alternatively
    obtain the result from the previous topological argument by using the family
    \(\mc K\) of compact sets in the constructed topology on
    \(\omega_k\) as our witness. Of course, every Lindel\"of set in
    a locally countable space is countable. Thus \(\mc K\)
    is cofinal in \([\omega_k]^\omega\)
    since for every countable set \(A\), \(\cl A\) is compact and countable.
    It is Kurepa since for every countable set \(A\), let (TODO)
  \end{proof}

  \begin{theorem}
    \(S'(\theta)\) holds whenever there exists a cofinal Kurepa family on \(\theta\).
  \end{theorem}

  \begin{proof}
    Let \(k<\omega\), and \(\mc K=\{K_\alpha:\alpha<\kappa\}\)
    be a cofinal Kurepa family on \(\theta\).
    We should define \(f_\alpha:K_\alpha\to\omega\) for each \(\alpha<\kappa\).

    Suppose we've defined pairwise almost compatible
    \(\{f_\beta:\beta<\alpha\}\). To define
    \(f_\alpha\), we first recall that \(\mc K\rest K_\alpha\) is countable,
    so we may choose \(\beta_n<\alpha\) for \(n<\omega\) such that
    \(
      \{K_\beta:\beta<\alpha\}\rest K_\alpha \setminus \{\emptyset\}
        =
      \{K_\alpha\cap K_{\beta_n}:n<\omega\}
    \).
    Let \(K_\alpha=\{\delta_{i,j}:i\leq\omega,j<w_i\}\) where
    \(w_i\leq\omega\) for each \(i\leq\omega\),
    \(
      K_\alpha\cap \left(K_{\beta_n}\setminus\bigcup_{m<n}K_{\beta_m}\right)
        =
      \{\delta_{n,j}:j<w_n\}
    \),
    and
    \(
      K_\alpha\setminus\bigcup_{n<\omega}K_{\beta_n}
        =
      \{
        \delta_{\omega,j}:j<w_\omega
      \}
    \).
    Then let \(f_\alpha(\delta_{n,j})=\max(n,f_{\beta_n}(\delta_{n,j}))\) for
    \(n<\omega\) and \(f_\alpha(\delta_{\omega,j})=j\) otherwise.

    We should show that \(f_\alpha\) is finite-to-one. Let \(n<\omega\).
    We need only worry about \(\delta_{m,j}\) for \(m\leq n\) since
    \(f_\alpha(\delta_{m,j})\geq m\). Since each
    \(f_{\beta_m}\) is finite-to-one, \(f_{\beta_m}(\delta_{m,j})\leq n\)
    for only finitely many \(j\). Thus \(f_\alpha\) maps to
    \(n\) only finitely often.

    We now want to demonstrate that \(f_\alpha\sim f_{\beta_n}\) for all
    \(n<\omega\). We again need only concern ourselves with \(\delta_{m,j}\) for
    \(m\leq n\) since otherwise \(\delta_{m,j}\not\in K_{\beta_n}\).
    For \(m=n\), we have
    \(f_\alpha(\delta_{n,j})=\max(n,f_{\beta_n}(\delta_{n,j}))\) which differs
    from \(f_{\beta_n}(\delta_{n,j})\) for only the finitely many \(j\) which
    are mapped below \(n\) by \(f_{\beta_n}\).
    For \(m<n\) and \(\delta_{m,j}\in K_{\beta_n}\),we have
    \(f_\alpha(\delta_{m,j})=\max(m,f_{\beta_m}(\delta_{m,j}))\) which can
    only differ
    from \(f_{\beta_n}(\delta_{m,j})\) for only the finitely many \(j\) which
    are mapped below \(m\) by \(f_{\beta_m}\) or the finitely many \(j\)
    for which the
    almost compatible \(f_{\beta_n}\sim f_{\beta_m}\) differ.
  \end{proof}

  \begin{corollary}
    \(S'(\omega_k)\) holds for all \(k<\omega\).
  \end{corollary}

  As noted in [TODO cite Dow],
  Jensen's one-gap two-cardinal theorem under \(V=L\) [TODO cite] can be used
  to show that there exist cofinal Kurepa families on every cardinal.

  \begin{corollary}[\(V=L\)]
    \(S'(\theta)\) holds for all cardinals.
  \end{corollary}

  In particular, \(S(\omega_2)\) fails under \(CH\), showing the two are
  unique. Actually, \(CH\) is not required to have \(S(\omega_2)\) fail.

  % TODO: read through Alan's proof and fit to rest of document
  \begin{theorem}
    Adding \(\omega_2\) Cohen reals to a model of \(CH\) forces
    \(\mf c=\omega_2\) and \(\neg S(\omega_2)\).
  \end{theorem}

  \begin{proof}
    TODO add Alan's proof
  \end{proof}
